---
title: "Descriptive Statistics Report"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(psych)
library(ggplot2)
library(tidyr)
library(reshape2)
library(ggpubr)
```

# 1. Introduction

This report presents descriptive statistics of the dataset, including central tendency, variability, distribution, and reliability of the measured items.

```{r}
pre_data <- pre$data |> mutate(time = "pre")
post_data <- post$data |> mutate(time = "post")

combined_ <- bind_rows(pre_data, post_data)

items <- names(pre$data)
combined <- combined_ |> mutate(across(all_of(items), ~ as.numeric(.)))
```

# 2. Descriptive Statistics
## 2.1 Overall Descriptives
```{r}
overall_prop <- combined |>
  summarise(across(all_of(items), ~ mean(.x, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "item", values_to = "proportion")

overall_prop |> print(n = Inf)
```

## 2.2 Descriptive by Group (Pre vs Post)
```{r}
prop_by_group <- combined |>
  group_by(time) |>
  summarise(across(all_of(items), ~ mean(.x, na.rm = TRUE))) |>
  pivot_longer(-time, names_to = "item", values_to = "proportion")

prop_by_group |> print(n = Inf)
```

# 3. Visualization
## 3.1 Proportion per Item (Pre vs Post)

```{r, fig.height=8}
combined_long <- combined_ |>
  pivot_longer(cols = all_of(items), names_to = "item", values_to = "value")

ggplot(combined_long, aes(x = value, fill = time)) +
  geom_bar(position = "dodge") +
  facet_wrap(~item, scales = "free") +
  labs(title = "Distribution per Item (Pre vs Post)") +
  theme_pubclean()
```

## 3.2 Boxplot (Pre vs Post)

```{r, fig.height=8}
diff_plot <- prop_by_group |>
  pivot_wider(names_from = time, values_from = proportion) |>
  mutate(diff = post - pre)

ggplot(diff_plot, aes(x = item, y = diff)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Change in Proportion (Post - Pre)",
    y = "Difference",
    x = "Item"
  ) +
  theme_pubclean()
```

# 4. Reliability Analysis
## 4.1 Cronbach’s Alpha (Pre-test)
```{r}
pre_poly <- psych::polychoric(pre_data[, items])
psych::alpha(pre_poly$rho)
```

## 4.2 Cronbach’s Alpha (Post-test)
```{r}
post_poly <- psych::polychoric(post_data[, items])
psych::alpha(post_poly$rho)
```

# 5. Correlation Matrix
```{r}
poly <- psych::polychoric(combined[, items])
cor_matrix <- poly$rho
round(cor_matrix, 2)
```

## Heatmap
```{r}
melted <- reshape2::melt(poly$rho)

ggplot(melted, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Heatmap") +
  theme_pubclean()
```

# 7. Interpretation
- Mean values indicate general response tendencies across items.
- Standard deviation reflects variability among participants.
- Skewness and kurtosis suggest whether distributions deviate from normality.
- Reliability (Cronbach’s alpha) indicates internal consistency.
- Correlation patterns help identify relationships among items.

# 8. Conclusion

The descriptive analysis provides a foundational understanding of the dataset, supporting further inferential analyses such as CFA and SEM.

---
