---
title: "Laporan Data Pretest"
email: "info@notivra.com"
author: "Notivra (www.notivra.com)"
date: "`r format(Sys.Date(), '%B %d, %Y')`"

lang: "en"
output:
  html_document:
    theme:
      version: 4
      bootswatch: flatly
      primary: "#ff9800"    # bright sunshine orange
      secondary: "#ffc107"  # warm yellow
    toc: true
    toc_float: true
    highlight: tango
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
    df_print: kable
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



```{r lavaan-expanded-interpretation, results='asis', echo=FALSE}
library(lavaan)
library(psych)

cat("## 🔍 Full CFA Interpretation\n\n")

# -------------------------------------------------------------
# 1 — FIT INDEX EXTRACTION
# -------------------------------------------------------------
fm <- fitMeasures(fit)

# regular
cfi   <- fm["cfi"];     tli   <- fm["tli"]
rmsea <- fm["rmsea"];   rmsea_l <- fm["rmsea.ci.lower"]; rmsea_h <- fm["rmsea.ci.upper"]
srmr  <- fm["srmr"]

# scaled / robust
cfi_s <- fm["cfi.scaled"];   tli_s <- fm["tli.scaled"]
rmsea_s <- fm["rmsea.scaled"]
rmsea_sl <- fm["rmsea.ci.lower.scaled"]; rmsea_sh <- fm["rmsea.ci.upper.scaled"]
chisq_s <- fm["chisq.scaled"]; df <- fm["df"]; p_s <- fm["pvalue.scaled"]

cat("###  Model Fit\n")
cat("Model fit lives or dies on three pillars: **CFI/TLI**, **RMSEA**, and **SRMR**.\n",
    "Because we use WLSMV, the *scaled* values carry the truth.\n\n")

cat("**Scaled CFI =", round(cfi_s,3), "** — higher is better (≥0.95 ideal).\n")
cat("**Scaled TLI =", round(tli_s,3), "** — reflects parsimony.\n")
cat("**Scaled RMSEA =", round(rmsea_s,3), "** (", round(rmsea_sl,3), "–", 
    round(rmsea_sh,3), ") — smaller is better (≤0.06 sweet).\n", sep="")
cat("**SRMR =", round(srmr,3), "** — closer to 0 is harmony.\n\n")

cat("**Scaled Chi-square:** χ²(", df, ") = ", round(chisq_s,3), 
    ", p = ", signif(p_s,3), ".\n\n", sep="")

# Short verdict
cat("**Verdict:**\n")
if (cfi_s >= .95 & rmsea_s <= .06) {
  cat("The model stands strong: factors align cleanly and the data respects the structure.\n\n")
} else if (cfi_s >= .90 & rmsea_s <= .08) {
  cat("The model holds reasonably well—some turbulence exists, but the backbone is intact.\n\n")
} else {
  cat("The model struggles. The structure may not reflect how respondents think.\n\n")
}

# -------------------------------------------------------------
# 2 — FACTOR-BY-FACTOR LOADING INTERPRETATION
# -------------------------------------------------------------
cat("###  Factor Loadings — Item Stories\n")
loads <- inspect(fit, "std")$lambda
loads <- round(loads, 3)

for (factor in rownames(loads)) {
  f_load <- loads[factor,]
  f_load <- f_load[f_load != 0]
  if (length(f_load) == 0) next
  
  cat("\n#### **", factor, "**\n", sep="")
  
  # Prepare item(label + number)
  labeled <- paste0(names(f_load), " (", f_load, ")")
  names(labeled) <- names(f_load)
  
  strong <- labeled[f_load >= .70]
  moderate <- labeled[f_load < .70 & f_load >= .40]
  weak <- labeled[f_load < .40]
  
  if (length(strong) > 0) {
    cat("*Strong anchors:* ", paste(strong, collapse=", "), "\n")
  }
  if (length(moderate) > 0) {
    cat("*Steady contributors:* ", paste(moderate, collapse=", "), "\n")
  }
  if (length(weak) > 0) {
    cat("*Shaky items:* ", paste(weak, collapse=", "), "\n")
  }
  
  cat("\n")
}

cat("\nA factor breathes well when most of its items load ≥0.50.  
Items below 0.40 speak with a trembling voice—they may need rewriting or removal.\n\n")

# -------------------------------------------------------------
# 3 — RELIABILITY (ω & α)
# -------------------------------------------------------------
cat("###  Reliability — Internal Strength\n")
data_num <- as.data.frame(lapply(items, function(x) as.numeric(as.character(x))))

# compute for each factor
lvs <- inspect(fit, "std")$lambda
factors <- rownames(lvs)

for (f in factors) {
  idx <- which(lvs[f,] != 0)
  if (length(idx) >= 3) {
    sub <- data_num[, idx, drop=FALSE]
    a <- psych::alpha(sub)$total$raw_alpha
    w <- psych::omega(sub)$omega.tot
    
    cat("**", f, ":**  α = ", round(a,3), 
        ", ω = ", round(w,3), "\n", sep="")
  }
}
cat("\nAlpha tells you how tightly items march together.  
Omega tells you whether the factor truly commands them.  
ω ≥ 0.70 usually means the construct is stable.\n\n")

# -------------------------------------------------------------
# 4 — MODIFICATION INDICES (OPTIONAL INSIGHTS)
# -------------------------------------------------------------
cat("###  Modification Indices — Where the Strain Lives\n")
mi <- modificationIndices(fit)
big <- mi[mi$mi > 10, ]

if (nrow(big)==0) {
  cat("The model whispers quietly—no large residual tensions.\n\n")
} else {
  cat("Items that want to talk to each other outside the script:\n\n")
  print(head(big[order(-big$mi), ], 10))
  cat("\nHigh MI suggests shared wording, overlapping meaning, or double-loading.\n\n")
}

cat("###  Closing Interpretation\n")
cat("The CFA is more than numbers—it's a map of how minds cluster ideas.  
When the scaled indices sing, when loadings rise steady, and reliability holds,  
you've carved a structure that reflects how respondents truly think.\n\n")
```


## Model Diagnostics and Interpretation

The CFA model converged under WLSMV, but several indicators suggest that the 
measurement model is not fully stable and may require revision.

###  Empty Tetrachoric Cells
The warning about *“two empty cells in a 2×2 table”* indicates that certain 
binary item pairs never co-occur in specific response combinations. This 
suggests severe imbalance in some items (e.g., nearly all 0s). This reduces 
tetrachoric reliability and may weaken factor estimation.

###  Near-Perfect Item Correlation
The message stating that *“the correlation between i14 and i12 is nearly 1.0”* 
implies that these two items are essentially duplicates. Such redundancy can 
destabilize the WLSMV estimator, inflate factor covariances, and produce 
ill-conditioned covariance matrices. Removing one of the items is recommended.

###  Non–Positive-Definite Latent Covariance Matrix
The latent covariance matrix was reported as *“not positive definite”*. This is 
a structural warning that the factors (C1–C4) may be too highly correlated or 
not empirically separable. It can also arise from duplicated or near-constant 
binary items compressing the factor space. This suggests the conceptual 
distinction between factors may not be strongly supported by the data.

###  Fit Index Pattern
The fit indices show mixed signals:

- **CFI/TLI (scaled):** high (>0.98) → good  
- **RMSEA (scaled):** ~0.041 → good  
- **SRMR:** 0.240 → poor  

The extremely high SRMR indicates that the model fails to reproduce many of the 
item correlations. In binary CFA, such patterns often arise from strong residual 
dependencies or misspecified factor structure.

###  Factor Loadings
Most items load significantly on their intended factors. However, a few items 
exhibit relatively low standardized loadings (<0.30), suggesting they may not 
measure the latent construct effectively.

## Summary
Although the model converged and several fit indices appear acceptable, the high 
SRMR, non–positive-definite latent covariance matrix, and near-duplicate items 
indicate that the current model is unstable. Reconsidering item inclusion, 
removing highly collinear items, and revising the factor structure are warranted 
before interpreting the results substantively.
