---
title: "`r params$title`"
output: pdf_document
params:
  title: "CFA Report (Pre–Post Analysis)"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(lavaan)
library(dplyr)
library(ggplot2)
library(semPlot)
```

# 1. Introduction

This study examines the measurement structure of the construct before and after intervention using Confirmatory Factor Analysis (CFA) with ordinal indicators (WLSMV estimator).

```{r}
pre_data <- pre$data |> mutate(time = "pre")
post_data <- post$data |> mutate(time = "post")

combined <- bind_rows(pre_data, post_data)

items <- c("i4", "i11", "i13", "i9", "i10", "i12", "i15")
```


# 2. Pre-test Measurement Model
```{r}
model_2f <- "
F1 =~ i4 + i11 + i13
F2 =~ i9 + i10 + i12 + i15
"

fit_pre <- cfa(
  model_2f,
  data = pre_data,
  estimator = "WLSMV",
  ordered = items
)

fitMeasures(fit_pre, c("cfi", "rmsea", "srmr"))
```

### Factor Loadings

```{r}
parameterEstimates(fit_pre) |>
  filter(op == "=~") |>
  select(Factor = lhs, Item = rhs, Loading = est, pvalue)
```


# 3. Post-test Measurement Model

## 3.1 Two-Factor Model
```{r}
fit_post_2f <- cfa(
  model_2f,
  data = post_data,
  estimator = "WLSMV",
  ordered = items
)

fitMeasures(fit_post_2f, c("cfi", "rmsea", "srmr"))
lavaan::inspect(fit_post_2f, "cor.lv")
```

## 3.2. One-Factor Model
```{r}
model_1f <- "
F =~ i4 + i11 + i13 + i9 + i10 + i12 + i15
"

fit_post_1f <- cfa(
  model_1f,
  data = post_data,
  estimator = "WLSMV",
  ordered = items
)

fitMeasures(fit_post_1f, c("cfi", "rmsea", "srmr"))
```


# 4. Model Comparison
```{r}
data.frame(
  Model = c("Pre (2F)", "Post (2F)", "Post (1F)"),
  CFI = c(
    fitMeasures(fit_pre, "cfi"),
    fitMeasures(fit_post_2f, "cfi"),
    fitMeasures(fit_post_1f, "cfi")
  ),
  RMSEA = c(
    fitMeasures(fit_pre, "rmsea"),
    fitMeasures(fit_post_2f, "rmsea"),
    fitMeasures(fit_post_1f, "rmsea")
  ),
  SRMR = c(
    fitMeasures(fit_pre, "srmr"),
    fitMeasures(fit_post_2f, "srmr"),
    fitMeasures(fit_post_1f, "srmr")
  )
)
```


# 5. Measurement Invariance (Exploratory)
```{r}
fit_config <- cfa(
  model_2f,
  data = combined,
  group = "time",
  estimator = "WLSMV",
  ordered = items
)

fit_metric <- cfa(
  model_2f,
  data = combined,
  group = "time",
  group.equal = "loadings",
  estimator = "WLSMV",
  ordered = items
)

data.frame(
  Model = c("Configural", "Metric"),
  CFI = c(
    fitMeasures(fit_config, "cfi"),
    fitMeasures(fit_metric, "cfi")
  ),
  RMSEA = c(
    fitMeasures(fit_config, "rmsea"),
    fitMeasures(fit_metric, "rmsea")
  ),
  SRMR = c(
    fitMeasures(fit_config, "srmr"),
    fitMeasures(fit_metric, "srmr")
  )
)
```


# 6. Visualization
### Pre-test Model

```{r}
semPaths(fit_pre, what = "std", layout = "tree")
```

### Post-test (Best Model)
```{r}
semPaths(fit_post_1f, what = "std", layout = "tree")
```


# 7. Interpretation
- The pre-test model supports a two-factor structure.
- The post-test two-factor model shows discriminant validity issues.
- The one-factor model provides a more stable representation.

Conclusion:
The intervention appears to have shifted the construct from multidimensional to unidimensional.


# 8. Conclusion

The findings suggest a structural transformation of the latent construct following intervention, indicating conceptual consolidation rather than mere score improvement.

---
